---
title: "un_datathon"
output:
  word_document: default
  html_document: default
date: "2023-11-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
install.packages('dplyr')
library(dplyr)
```

```{r}
internet_use <- UNdata_Export_20231104_183511233
internet_use = internet_use %>% slice(1:25)
mobile_use <- read.csv("mobile.csv")
mobile_use = mobile_use %>% slice(1:25)
tele_use <- read.csv("tele.csv")
tele_use = tele_use %>% slice(1:25)

library("ggplot2")
ggplot(data = internet_use, aes(x=Year, y=Value)) + 
  geom_line() +
  theme_minimal() +
  ggtitle("Percentage of Individuals in Nigeria Using the Internet (1990-2014)") +
  theme(panel.grid = element_blank()) + 
  ylab("Percentage")  
```

```{r}
jpeg(filename = "internetuse.jpeg")
library("ggplot2")
ggplot(data = internet_use, aes(x=Year, y=Value)) + 
  geom_line() +
  theme_minimal() +
  ggtitle("Percentage of Individuals in Nigeria Using the Internet (1990-2014)") +
  theme(panel.grid = element_blank()) + 
  ylab("Percentage")  
dev.off()
```

```{r}
schl_connect <- read_csv("schl_connect.csv")
jpeg(filename = "internet_use_schl.jpeg")
barplot(schl_connect$Perc, names.arg = schl_connect$Education_Level, col = "darkblue", xlab = "Education Level", ylab = "Percentage", main = "Percentage of Children with\nInternet Connection at home by Education Level")
dev.off()
```

sectc1_plantingw4: Data collected through Post Planting Community Questionnaire, Section C1 (Respondent Characteristics)
sect4b_plantingw4: Data collected through Post Planting Household Questionnaire, Section 4B (ICT - Mobile Phone Banking)

```{r}
char <- read_csv("sectc1_plantingw4.csv")
ict <- read_csv("sect4b_plantingw4.csv")
```
```{r}
# subset to columns of interest (char)
char = subset(char, select = -c(10:13))
char <- char %>% rename(urban = sector,
                        sex = c01q02,
                        age = c01q03,
                        edu = c01q06)

# subset to columns of interest (ict)
ict = subset(ict, select = -c(6,8,10,19,20))
ict <- ict %>% rename(mobilephone = s4bq8,
                      internet = s4bq14,
                      personal_device = s4bq15__1,
                      other_household = s4bq15__2,
                      relative_friend = s4bq15__3,
                      workplace = s4bq15__4,
                      cafe = s4bq15__5,
                      hotspot = s4bq15__6)

# merge dataframes by lga
dta <- merge(char, ict, by = "lga")

# delete irrelevant columns again
dta = subset(dta, select = -c(11:14,7))

# make a copy
d <- dta

# cleaning education categories
d$edu <- ifelse(d$edu == 0,0, d$edu)
d$edu <- ifelse(d$edu %in% c(1,2),1, d$edu)
d$edu <- ifelse(d$edu %in% c(11,12,13,14,15,16), 2, d$edu)
d$edu <- ifelse(d$edu %in% c(21,22,23), 3, d$edu)
d$edu <- ifelse(d$edu %in% c(24,25,26), 4, d$edu)
d$edu <- ifelse(d$edu == 27, 5, d$edu)
d$edu <- ifelse(d$edu == 28, 6, d$edu)
d$edu <- ifelse(d$edu %in% c(31,33,34,41,42,43, 51, 52, 61, 321, 322), 7, d$edu)

# cleaning mobile and internet access variable
d$mobilephone <- ifelse(is.na(d$mobilephone), NA, ifelse(d$mobilephone == 2, 0, d$mobilephone))
d$internet <- ifelse(is.na(d$internet), NA, ifelse(d$internet == 2, 0, d$internet))

# cleaning sector variable
d$urban <- ifelse(is.na(d$urban), NA, ifelse(d$urban == 2, 0, d$urban))
```

```{r}
# Calculate the percentage of individuals with access to mobile phones
mobile_access <- table(d$mobilephone)
mobile_percentage <- (mobile_access[["1"]] / sum(mobile_access)) * 100

# Calculate the percentage of individuals with internet access
internet_access <- table(d$internet)
internet_percentage <- (internet_access[["1"]] / sum(internet_access)) * 100

# Print the results
cat("Percentage with Mobile Phone Access: ", mobile_percentage, "%\n")
cat("Percentage with Internet Access: ", internet_percentage, "%\n")

```

```{r}
library(ggplot2)
# Assuming "education" is your education variable and "mobile_phone" is the mobile phone ownership variable (1 for owning)
d$edu <- factor(d$edu, levels = 0:8)
ggplot(na.omit(d), aes(x = edu, fill = factor(mobilephone))) +
  geom_bar() +
  labs(title = "Number of People Owning Mobile Phones by Education Level", x = "Education Level", y = "Count") +
  scale_fill_manual(values = c("1" = "royalblue"), name = "Has access to a mobile phone 
                    0: No education
                    1:Nursery
                    2: Primary
                    3: Junior Secondary
                    4: Senior Secondary
                    5: Lower 6
                    6: Upper 6
                    7: Higher Education") +
  theme_classic() 
```

```{r}
d1 <- d %>% filter(!is.na(edu) & !is.na(mobilephone))

ggplot(d1, aes(x = edu, fill = factor(internet))) +
  geom_bar() +
  labs(title = "Number of People with Access to Internet by Education Level", x = "Education Level", y = "Count") +
  scale_fill_manual(values = c("1" = "royalblue", "0" = "darkred"), name = "Has access to internet 
                    0: No education
                    1:Nursery
                    2: Primary
                    3: Junior Secondary
                    4: Senior Secondary
                    5: Lower 6
                    6: Upper 6
                    7: Higher Education") +
  theme_classic()
```

```{r}
library(dplyr)

# Calculate the percentage of internet access by urban/rural population
internet_percentage <- d %>%
  group_by(urban) %>%
  summarize(internet_access = sum(!is.na(internet) & internet == "1") / sum(!is.na(internet)))


# Calculate the percentage of mobile phone access by urban/rural population
mobile_percentage <- d %>%
  group_by(urban) %>%
  summarize(mobile_phone_access = sum(!is.na(mobilephone) & mobilephone == "1") / sum(!is.na(mobilephone)))
```

```{r}
internet_percentage
```
Shows that there is a drastic difference in access to internet among urban populations (1) compared to rural (0)

```{r}
mobile_percentage
```
Interestingly, not as drastic for mobile phone access. 

Delving into where people commonly get their internet access from:
```{r}
# Update the loc_access variable based on the binary variables
d$loc_access[d$personal_device == 1] <- 1
d$loc_access[d$other_household == 1] <- 2
d$loc_access[d$relative_friend == 1] <- 3
d$loc_access[d$workplace == 1] <- 4
d$loc_access[d$cafe == 1] <- 5
d$loc_access[d$hotspot == 1] <- 6

d1 <- na.omit(d)
# Assuming you have a dataset named "d1"
# Assuming you have a dataset named "d1"
library(ggplot2)

ggplot(d1, aes(x = factor(edu), fill = factor(loc_access))) +
  geom_bar() +
  labs(title = "Accessing Internet by Education Level",
       x = "Education Level", y = "Counts") +
  scale_fill_brewer(name = "Where do they access the internet?", palette = "Spectral", 
                    labels = c("1: Personal Device", 
                               "2: Household", 
                               "3: Relative/Friend",
                               "4: Workplace",
                               "5: Cafe",
                               "6: Hotspot")) +
  theme(legend.position = "top") +
  theme_classic() +
  coord_flip()

```
```{r}
write.csv(d, file = "nigeria_hsehold.csv", row.names = FALSE)
```
